---
author: Sam White
toc-title: Contents
toc-depth: 5
toc-location: left
title: Genome Assembly - C.rogercresseyi Macho with PacBio HiFi Reads Using hifiasm on Hyak
date: '2026-05-19'
draft: false
engine: knitr
execute:
  eval: false
categories: 
  - "2026"
  - "sea lice"
  - "hyak"
  - "Caligus rogercresseyi"
  - "PacBio"
  - "hifiasm"
  - "genome assembly"
---
# INTRO

Steven [asked the I assemble sea lice PacBio data using hifiasm](https://github.com/RobertsLab/resources/issues/2421) (GitHub Issue).

I assembled the PacBio HiFi reads for the _C. rogercresseyi_ Macho sample using hifiasm on Hyak using our Apptainer image: `srlab-R4.4-bioinformatics-container-c3d3116.sif`. Assembly took ~19hrs, which was significantly longer then [the Hembra assembly (1.5hrs)](../2026-05-19-Genome-Assembly---C.rogercresseyi-Hembra-with-PacBio-HiFi-Reads-Using-hifiasm-on-Hyak/index.qmd) (Notebook entry).

Now that both assemblies are complete, I'll compare the two assemblies using QUAST.

# RESULTS

Output directory:

- [https://gannet.fish.washington.edu/Atumefaciens/20260519-crog-macho-genome_assembly-hifiasm/](https://gannet.fish.washington.edu/Atumefaciens/20260519-crog-macho-genome_assembly-hifiasm/)

Assembly FastA:

- [https://gannet.fish.washington.edu/Atumefaciens/20260519-crog-macho-genome_assembly-hifiasm/crog-macho-pb-hifiasm-assembly.fa](https://gannet.fish.washington.edu/Atumefaciens/20260519-crog-macho-genome_assembly-hifiasm/crog-macho-pb-hifiasm-assembly.fa) (1009MB)



# VARIABLES
```{r set-variables}
# DIRECTORIES
data_dir <- "/mmfs1/gscratch/scrubbed/samwhite/data/C_rogercresseyi/HIFI_Caligus/Macho_01_cell1/"
output_dir <- "/mmfs1/gscratch/scrubbed/samwhite/outputs/20260519-crog-macho-genome_assembly-hifiasm/"

# FILES
hifiasm_asm <- "crog-macho-pb-hifiasm-assembly.asm"
hifiasm_primary_graph <- "crog-macho-pb-hifiasm-assembly.asm.bp.p_ctg.gfa"
hifiasm_assembly_fasta <- "crog-macho-pb-hifiasm-assembly.fa"
hifiasm_assembly_log <- "crog-macho-pb-hifiasm-assembly.log"

# PROGRAMS
gfatools <- c("/srlab/programs/gfatools/gfatools")
hifiasm <- c("/srlab/programs/miniforge3-24.7.1-0/envs/hifiasm-0.25.0_env/bin/hifiasm")


# SETTINGS
threads <- "50"


# Export these as environment variables for bash chunks.
Sys.setenv(
  data_dir = data_dir,
  gfatools = gfatools,
  hifiasm = hifiasm,
  hifiasm_asm = hifiasm_asm,
  hifiasm_assembly_fasta = hifiasm_assembly_fasta,
  hifiasm_assembly_log = hifiasm_assembly_log,
  hifiasm_primary_graph = hifiasm_primary_graph,
  output_dir = output_dir,
  threads = threads
)
```


# HIFIASM
```{r engine='bash'}
# Make output directory, if it doesn't exist
mkdir --parents "${output_dir}"

cd "${output_dir}"

# Create an array of .fastq.gz files
fastqs=(${data_dir}/*.fastq.gz)

# Print the result (optional, for verification)
## newline-delimited format
echo "List of FastQs used for assembly:"

printf "%s\n" "${fastqs[@]}" \
| tee input_fastqs.txt

# Run hifiasm assembly
"${hifiasm}" \
-o "${hifiasm_asm}" \
-t "${threads}" \
"${fastqs[@]}" \
> "${hifiasm_assembly_log}" 2>&1
```

# 3 GFATOOLS

## 3.1 Primary contigs to FastA

This will extract a FastA file from the graph files output by hifiasm.

```{r engine='bash'}
cd "${output_dir}"

"${gfatools}" gfa2fa \
"${hifiasm_primary_graph}" \
> "${hifiasm_assembly_fasta}"

samtools faidx "${hifiasm_assembly_fasta}"
```


# CHECKSUMS

```{r checksums, engine='bash'}
cd "${output_dir}"

for fasta in *.fa
do
  md5sum "${fasta}" \
  | tee "${fasta}".md5
done
```