---
author: Sam White
toc-title: Contents
toc-depth: 5
toc-location: left
title: Genome Comparisons - C.rogercresseyi Hembra and Macho PacBio hifiasm Assemblies Using QUAST on Hyak
date: '2026-05-20'
draft: true
engine: knitr
categories: 
  - "2026"
  - "hyak"
  - "Caligus rogercresseyi"
  - "QUAST"
  - "sea lice"
---
# INTRO
I previously assembled two _C. rogercresseyi_ genomes (Hembra and Macho) using hifiasm on Hyak:

- [2026-05-19-Genome-Assembly---C.rogercresseyi-Hembra-with-PacBio-HiFi-Reads-Using-hifiasm-on-Hyak](../2026-05-19-Genome-Assembly---C.rogercresseyi-Hembra-with-PacBio-HiFi-Reads-Using-hifiasm-on-Hyak/index.qmd) (Notebook entry)

- [2026-05-19-Genome-Assembly---C.rogercresseyi-Macho-with-PacBio-HiFi-Reads-Using-hifiasm-on-Hyak](../2026-05-19-Genome-Assembly---C.rogercresseyi-Macho-with-PacBio-HiFi-Reads-Using-hifiasm-on-Hyak/index.qmd) (Notebook entry)

Used [QUAST](https://quast.sourceforge.net/docs/manual.html) (SourceForge) to evaluate the two sets of PacBio genome assemblies produced with [hifiasm](https://github.com/chhylp123/hifiasm) (GitHub). QUAST was run on Hyak using our Apptainer image: `srlab-R4.4-bioinformatics-container-c3d3116.sif`.

# RESULTS

Output directory:

- [https://gannet.fish.washington.edu/Atumefaciens/20260520-crog-quast/](https://gannet.fish.washington.edu/Atumefaciens/20260520-crog-quast/)


Code is below.

---


# SETUP
## Libraries
```{r setup}
library(knitr)
library(tidyverse)
```

    ## ── Attaching core tidyverse packages ──────────────────────── tidyverse 2.0.0 ──
    ## ✔ dplyr     1.1.4     ✔ readr     2.1.5
    ## ✔ forcats   1.0.0     ✔ stringr   1.5.1
    ## ✔ ggplot2   3.5.2     ✔ tibble    3.2.1
    ## ✔ lubridate 1.9.4     ✔ tidyr     1.3.1
    ## ✔ purrr     1.0.4     
    ## ── Conflicts ────────────────────────────────────────── tidyverse_conflicts() ──
    ## ✖ dplyr::filter() masks stats::filter()
    ## ✖ dplyr::lag()    masks stats::lag()
    ## ℹ Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
## Kniter Options
```{r set-knitr-options}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
    eval = FALSE,        # Evaluate code chunks
    results = "hold",   # Hold outputs and show them after the full code chunk
  warning = FALSE,     # Hide warnings
        collapse = FALSE,    # Keep code and output in separate blocks
        warning = FALSE,     # Hide warnings
        message = FALSE,     # Hide messages
        comment = "##"      # Prefix output lines with '##' so output is visually distinct
)
```

## Variables
```{r set-variables}
# OUTPUT DIRECTORY
output_dir <- "/mmfs1/gscratch/scrubbed/samwhite/outputs/20260520-crog-quast"

# INPUT FILES
hifiasm_hembra_fasta <- "/mmfs1/gscratch/scrubbed/samwhite/outputs/20260519-crog-hembra-genome_assembly-hifiasm/crog-hembra-pb-hifiasm-assembly.fa"
hifiasm_macho_fasta <- "/mmfs1/gscratch/scrubbed/samwhite/outputs/20260519-crog-macho-genome_assembly-hifiasm/crog-macho-pb-hifiasm-assembly.fa"

# SETTINGS
threads <- "40"


# Export these as environment variables for bash chunks.
Sys.setenv(
  hifiasm_hembra_fasta = hifiasm_hembra_fasta,
  hifiasm_macho_fasta = hifiasm_macho_fasta,
  output_dir = output_dir,
  threads = threads
)
```

# QUAST

```{r QUAST, engine=bash}
# Make output directory, if it doesn't exist
mkdir --parents "${output_dir}"

quast.py \
"${hifiasm_hembra_fasta}" \
"${hifiasm_macho_fasta}" \
--output-dir "${output_dir}" \
--threads "${threads}" \
--labels "hembra_hifiasm, macho_hifiasm"
```