---
author: Sam White
toc-title: Contents
toc-depth: 5
toc-location: left
title: Glycogen Assay - Glycogen-Glo Dilution Testing with M.gigas Ctenidia Homogenate
date: '2026-01-12'
draft: false
engine: knitr
categories: 
  - "2026"
  - "ctenidia"
  - "Glycogen-Glo"
  - "glycogen"
  - "Pacific oyster"
  - "Crassostrea gigas"
  - "Magallana gigas"
  - "plate reader"
  - "Gen 5"
  - "Synergy HTX"
---
# INTRO

This was the initial testing of the [Glycogen-Glo Assay (Promega)](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/Promega_Glycogen_Glo_Assay.pdf) (GitHub; PDF) using _M.gigas_ (Pacific oyster) ctenidia homegenate from [20251210](../../2025/2025-12-10-Tissue-Collection-and-Homogenization---M.gigas-Ctenidia-for-Glycogen-Glo-Assay/index.qmd) (notebook entry).

The idea was to identify a range (i.e. a very rough min/max) of starting tissues amounts that will fall within the range of the Glycogen-Glo standard curves. Additionally, this was an initial walk-through of using this kit to help figure out a general protocol that identified any tips/tricks/pitfalls to aid in developing a Standard Operating Protocol (SOP) for the [Roberts Lab Handbook](https://robertslab.github.io/resources/).

The assay will assess both glycogen _and_ glucose content. The way the assay works is to convert any glycogen to glucose. As such, glucose levels also need to be assessed for each sample, in order to _subtract_ that background glucose and get accurate glycogen quantification.

Overall, this process did not go very smoothly, but did highlight several challenges that can be addressed in future runs. Granted, I feel like most of this was related to my personal screw ups, but these will still serve as good guidance on improving this protocol for others.

# MATERIALS & METHODS

## Sample Selection

Sample [A2](../../2025/2025-12-10-Tissue-Collection-and-Homogenization---M.gigas-Ctenidia-for-Glycogen-Glo-Assay/index.qmd) was selected due to it's large tissue amount (50mg). This would allow us to perform serial dilutions representing different starting tissue amounts (roughly).

## Assay Preparation

### Samples and Standard Curves

Per the manufacturer's recommendations, all reagents were thawed. After thawing, the Glucoamylase Buffer and Luciferin Detection Solution were kept at room temperature, while the remaining components, including the A2 sample, were kept on ice.

The assay was set up to measure the following:

- A2 _GLYCOGEN_ content in the following equivalent "tissue" concentrations:

    - Not diluted (ND), 25mg, 2.5mg, 0.25mg, 0.025mg, 0mg

- A2 _GLUCOSE_ content in the following equivalent "tissue" concentrations:

    - Not diluted (ND), 25mg, 2.5mg, 0.25mg, 0.025mg, 0mg

- GLYCOGEN Standard curve:

   - 20ug/uL, 2ug/mL, 0.2ug/uL, 0.02ug/uL, 0ug/uL

- GLUCOSE Standard curve:

    - 100mM, 10mM, 1mM, 01.mM, 0mM

All dilutions were performed using a 2:1:1 mixture of PBS, 0.3N HCl, and TRIS Buffer (the latter two are supplied with the kit).

The A2 dilutions were performed in duplicate; primarily due to my poor calculations on the amount of the dilution buffer needed. Ugh.

::: {.callout-warning}
The 25mg tissue equivalents were highly inconsistent due to pipetting issues and should be disregarded! This was caused by insoluble debris.

For future sample handling, sample homogenate should be thawed, vortexed, and centrifuged prior to pipetting to avoid this issue.
:::

The standard curves were each performed in triplicate.


## Assay

Samples and standard curves were prepared according to the manufacturer's instructions. Theh plate used was a Costar 96-well, white, opaque, flat-bottom. 

Briefly, 25uL of sample/standard was added to wells (see the [plate layout below](#plate-layout)). For those samples/standards being assessed for glycogen, 25uL of glucoamylase digestion solution was added. For those samples/standards being assessed for just glucose, 25uL of _only_ Glucoamylase Buffer was added. The plate was loosely covered with aluminum foil to prevent light/debris from getting in the wells, gently shaken by hand for 60s, and incubated at room temp for 1hr. All wells were loaded individually.

After the initial incubation period, 50uL of glucose detection reagent was added to _ALL_ wells using an Eppendorf repeater pipette (Repeater M4). The plate was loosely covered with aluminum foil to prevent light/debris from getting in the wells, gently shaken by hand for 60s, and incubated at room temp for 1hr.

After incubating, luminescence was measured on the Roberts Lab Synergy HTX (Agilent) with the Gen5 software (Agilent).

## Plate layout

`blank` = Blank measurements (buffer only - no sample/standard added)

`SPL1` = Glycogen measurements of A2 homogenate.

`SPL2` = Glucose measurements of A2 homogenate.

`STDGly` = Glycogen standard curve measurement.

`STDGlu` = Glucose standard curve measurement.


|   | 1       | 2       | 3       | 4       | 5       | 6     | 7 | 8       | 9       | 10      | 11      | 12      |          |
|---|---------|---------|---------|---------|---------|-------|---|---------|---------|---------|---------|---------|----------|
| A | SPL1:1  | SPL1:2  | SPL1:3  | SPL1:4  | SPL1:5  | blank |   |         |         |         |         |         | Well ID  |
|   | 50      | 25      | 2.5     | 0.25    | 0.025   |       |   |         |         |         |         |         | Conc/Dil |
| B | SPL1:1  | SPL1:2  | SPL1:3  | SPL1:4  | SPL1:5  | blank |   |         |         |         |         |         | Well ID  |
|   | 50      | 25      | 2.5     | 0.25    | 0.025   |       |   |         |         |         |         |         | Conc/Dil |
| C | SPL2:1  | SPL2:2  | SPL2:3  | SPL2:4  | SPL2:5  | blank |   |         |         |         |         |         | Well ID  |
|   | 50      | 25      | 2.5     | 0.25    | 0.025   |       |   |         |         |         |         |         | Conc/Dil |
| D | SPL2:1  | SPL2:2  | SPL2:3  | SPL2:4  | SPL2:5  | blank |   |         |         |         |         |         | Well ID  |
|   | 50      | 25      | 2.5     | 0.25    | 0.025   |       |   |         |         |         |         |         | Conc/Dil |
| E | STDGly1 | STDGly2 | STDGly3 | STDGly4 | STDGly1 |       |   | STDGlu1 | STDGlu2 | STDGlu3 | STDGlu4 | STDGlu5 | Well ID  |
|   | 20      | 2       | 0.2     | 0.02    | 0       |       |   | 100     | 10      | 1       | 0.1     | 0       | Conc/Dil |
| F | STDGly1 | STDGly2 | STDGly3 | STDGly4 | STDGly1 |       |   | STDGlu1 | STDGlu2 | STDGlu3 | STDGlu4 | STDGlu5 | Well ID  |
|   | 20      | 2       | 0.2     | 0.02    | 0       |       |   | 100     | 10      | 1       | 0.1     | 0       | Conc/Dil |
| G | STDGly1 | STDGly2 | STDGly3 | STDGly4 | STDGly1 |       |   | STDGlu1 | STDGlu2 | STDGlu3 | STDGlu4 | STDGlu5 | Well ID  |
|   | 20      | 2       | 0.2     | 0.02    | 0       |       |   | 100     | 10      | 1       | 0.1     | 0       | Conc/Dil |
| H |         |         |         |         |         |       |   |         |         |         |         |         | Well ID  |
|   |         |         |         |         |         |       |   |         |         |         |         |         | Conc/Dil |

# RESULTS



## Raw fluorescence

|   | 1      | 2      | 3     | 4     | 5     | 6    | 7 | 8      | 9     | 10   | 11  | 12  |
|---|--------|--------|-------|-------|-------|------|---|--------|-------|------|-----|-----|
| A | 185099 | 174562 | 61687 | 7654  | 1762  | 813  |   |        |       |      |     |     |
| B | 166410 | 181069 | 83790 | 19910 | 10704 | 3187 |   |        |       |      |     |     |
| C | 843    | 744    | 421   | 237   | 227   | 689  |   |        |       |      |     |     |
| D | 396    | 230    | 261   | 257   | 221   | 234  |   |        |       |      |     |     |
| E | 107660 | 9963   | 1968  | 1277  | 1164  |      |   | 117932 | 12389 | 1703 | 673 | 606 |
| F | 110174 | 10220  | 1875  | 1256  | 737   |      |   | 122103 | 12418 | 1748 | 638 | 499 |
| G | 102674 | 10001  | 1815  | 1126  | 963   |      |   | 121056 | 12415 | 1629 | 480 | 356 |
| H |        |        |       |       |       |      |   |        |       |      |     |     |

## Glycogen Standard Curve

```{r glycogen-std-curve}
#| results: hold

library(ggplot2)

# Glycogen standard curve data from rows E, F, G, columns 1-5
# Concentrations in ug/uL
concentrations <- c(20, 2, 0.2, 0.02, 0)

# Luminescence values for each replicate
row_E <- c(107660, 9963, 1968, 1277, 1164)
row_F <- c(110174, 10220, 1875, 1256, 737)
row_G <- c(102674, 10001, 1815, 1126, 963)

# Calculate mean and standard error for each concentration
means <- numeric(5)
std_errors <- numeric(5)
std_devs <- numeric(5)

for (i in 1:5) {
  values <- c(row_E[i], row_F[i], row_G[i])
  means[i] <- mean(values)
  std_devs[i] <- sd(values)
  std_errors[i] <- sd(values) / sqrt(length(values))
}

# Create data frame for plotting
summary_data <- data.frame(
  concentration = concentrations,
  mean_luminescence = means,
  se = std_errors,
  sd = std_devs,
  cv = (std_devs / means) * 100
)

# Calculate linear regression and R-squared
lm_model <- lm(mean_luminescence ~ concentration, data = summary_data)
r_squared <- summary(lm_model)$r.squared
slope <- coef(lm_model)[2]
intercept <- coef(lm_model)[1]

# Sample data from row A (columns 1, 3, 4, 5, 6)
sample_data <- data.frame(
  tissue_mg = c(50, 2.5, 0.25, 0.025, 0),
  luminescence = c(185099, 61687, 7654, 1762, 813),
  label = c("50 mg", "2.5 mg", "0.25 mg", "0.025 mg", "0 mg")
)

# Create the plot
plot <- ggplot(summary_data, aes(x = concentration, y = mean_luminescence)) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = mean_luminescence - se, ymax = mean_luminescence + se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +
  geom_point(data = sample_data, aes(x = tissue_mg, y = luminescence,
             color = label, shape = label), size = 4) +
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "50 mg" = "darkred",
                                "2.5 mg" = "darkorange",
                                "0.25 mg" = "darkgreen",
                                "0.025 mg" = "purple",
                                "0 mg" = "brown")) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                "50 mg" = 17,
                                "2.5 mg" = 15,
                                "0.25 mg" = 18,
                                "0.025 mg" = 8,
                                "0 mg" = 4)) +
  annotate("label", x = max(summary_data$concentration) * 1.2, 
           y = max(summary_data$mean_luminescence) * 0.15,
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", slope, intercept, r_squared),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glycogen Standard Curve",
    x = "Glycogen Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = concentrations) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )

# Save the plot
ggsave("glycogen_standard_curve.png", plot, width = 10, height = 6, dpi = 300)

# Display the plot
plot

# Print summary statistics
cat("Glycogen Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(summary_data)) {
  cat(sprintf("Concentration: %g µg/µL\n", summary_data$concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n", summary_data$mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", summary_data$se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", summary_data$cv[i]))
}
```

## Glucose Standard Curve

```{r glucose-std-curve}
#| results: hold

library(ggplot2)

# Glucose standard curve data from rows E, F, G, columns 8-12
# Concentrations in mM
concentrations_glu <- c(100, 10, 1, 0.1, 0)

# Luminescence values for each replicate
row_E_glu <- c(117932, 12389, 1703, 673, 606)
row_F_glu <- c(122103, 12418, 1748, 638, 499)
row_G_glu <- c(121056, 12415, 1629, 480, 356)

# Calculate mean and standard error for each concentration
means_glu <- numeric(5)
std_errors_glu <- numeric(5)
std_devs_glu <- numeric(5)

for (i in 1:5) {
  values <- c(row_E_glu[i], row_F_glu[i], row_G_glu[i])
  means_glu[i] <- mean(values)
  std_devs_glu[i] <- sd(values)
  std_errors_glu[i] <- sd(values) / sqrt(length(values))
}

# Create data frame for plotting
summary_data_glu <- data.frame(
  concentration = concentrations_glu,
  mean_luminescence = means_glu,
  se = std_errors_glu,
  sd = std_devs_glu,
  cv = (std_devs_glu / means_glu) * 100
)

# Calculate linear regression and R-squared
lm_model_glu <- lm(mean_luminescence ~ concentration, data = summary_data_glu)
r_squared_glu <- summary(lm_model_glu)$r.squared
slope_glu <- coef(lm_model_glu)[2]
intercept_glu <- coef(lm_model_glu)[1]

# Sample data from row C (columns 1, 3, 4, 5, 6)
sample_data_glu <- data.frame(
  tissue_mg = c(50, 2.5, 0.25, 0.025, 0),
  luminescence = c(843, 421, 237, 227, 689),
  label = c("50 mg", "2.5 mg", "0.25 mg", "0.025 mg", "0 mg")
)

# Create the plot
plot_glu <- ggplot(summary_data_glu, aes(x = concentration, y = mean_luminescence)) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = mean_luminescence - se, ymax = mean_luminescence + se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +
  geom_point(data = sample_data_glu, aes(x = tissue_mg, y = luminescence,
             color = label, shape = label), size = 4) +
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "50 mg" = "darkred",
                                "2.5 mg" = "darkorange",
                                "0.25 mg" = "darkgreen",
                                "0.025 mg" = "purple",
                                "0 mg" = "brown")) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                "50 mg" = 17,
                                "2.5 mg" = 15,
                                "0.25 mg" = 18,
                                "0.025 mg" = 8,
                                "0 mg" = 4)) +
  annotate("label", x = max(summary_data_glu$concentration) * 0.7, 
           y = max(summary_data_glu$mean_luminescence) * 0.2,
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", slope_glu, intercept_glu, r_squared_glu),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glucose Standard Curve",
    x = "Glucose Concentration (mM)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = concentrations_glu) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )

# Save the plot
ggsave("glucose-std-curve.png", plot_glu, width = 10, height = 6, dpi = 300)

# Display the plot
plot_glu

# Print summary statistics
cat("Glucose Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(summary_data_glu)) {
  cat(sprintf("Concentration: %g mM\n", summary_data_glu$concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n", summary_data_glu$mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", summary_data_glu$se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", summary_data_glu$cv[i]))
}
```
## Sample A2 Quantification

```{r sample-quantification}
#| results: hold

# Function to calculate concentration from luminescence using linear model
# Returns concentration or "OUT OF RANGE" message
calculate_concentration <- function(luminescence, slope, intercept, min_conc, max_conc, units) {
  # Calculate concentration from y = mx + b -> x = (y - b) / m
  concentration <- (luminescence - intercept) / slope
  
  # Check if within range
  if (concentration < min_conc) {
    return("OUT OF RANGE - TOO LOW")
  } else if (concentration > max_conc) {
    return("OUT OF RANGE - TOO HIGH")
  } else {
    return(sprintf("%.4f %s", concentration, units))
  }
}

# Glycogen calculations
cat("GLYCOGEN ASSAY RESULTS (Sample A2):\n")
cat(rep("=", 70), "\n", sep = "")
cat(sprintf("%-15s %-20s %s\n", "Tissue Amount", "Luminescence", "Calculated Glycogen"))
cat(rep("-", 70), "\n", sep = "")

glycogen_min <- min(concentrations)
glycogen_max <- max(concentrations)

for (i in 1:nrow(sample_data)) {
  result <- calculate_concentration(
    sample_data$luminescence[i],
    slope,
    intercept,
    glycogen_min,
    glycogen_max,
    "µg/µL"
  )
  cat(sprintf("%-15s %-20.0f %s\n", 
              sample_data$label[i], 
              sample_data$luminescence[i], 
              result))
}

cat("\n\n")

# Glucose calculations
cat("GLUCOSE ASSAY RESULTS (Sample A2):\n")
cat(rep("=", 70), "\n", sep = "")
cat(sprintf("%-15s %-20s %s\n", "Tissue Amount", "Luminescence", "Calculated Glucose"))
cat(rep("-", 70), "\n", sep = "")

glucose_min <- min(concentrations_glu)
glucose_max <- max(concentrations_glu)

for (i in 1:nrow(sample_data_glu)) {
  result <- calculate_concentration(
    sample_data_glu$luminescence[i],
    slope_glu,
    intercept_glu,
    glucose_min,
    glucose_max,
    "mM"
  )
  cat(sprintf("%-15s %-20.0f %s\n", 
              sample_data_glu$label[i], 
              sample_data_glu$luminescence[i], 
              result))
}
```


# DISCUSSION

## Recommended Tissue Amounts


```{r tissue-range-recommendation}
#| results: hold

# Calculate the luminescence range for the glycogen standard curve
std_curve_min_lum <- min(summary_data$mean_luminescence)
std_curve_max_lum <- max(summary_data$mean_luminescence)

cat("GLYCOGEN STANDARD CURVE RANGE:\n")
cat(rep("=", 60), "\n", sep = "")
cat(sprintf("Minimum luminescence: %.2f (at %g µg/µL glycogen)\n", 
            std_curve_min_lum, min(concentrations)))
cat(sprintf("Maximum luminescence: %.2f (at %g µg/µL glycogen)\n\n", 
            std_curve_max_lum, max(concentrations)))

# Use the linear regression to predict tissue amounts that would fall within the standard curve range
# Using the equation: tissue_amount = (luminescence - intercept) / slope
predicted_min_tissue <- (std_curve_min_lum - intercept) / slope
predicted_max_tissue <- (std_curve_max_lum - intercept) / slope

cat("PREDICTED TISSUE AMOUNTS USING STANDARD CURVE:\n")
cat(rep("=", 60), "\n", sep = "")
cat(sprintf("Equation: y = %.2fx + %.2f\n\n", slope, intercept))
cat(sprintf("Minimum tissue amount: %.4f mg\n", predicted_min_tissue))
cat(sprintf("  (produces luminescence: %.2f)\n\n", std_curve_min_lum))
cat(sprintf("Maximum tissue amount: %.2f mg\n", predicted_max_tissue))
cat(sprintf("  (produces luminescence: %.2f)\n\n", std_curve_max_lum))

cat(rep("=", 60), "\n", sep = "")
cat("RECOMMENDED TISSUE RANGE FOR FUTURE ASSAYS:\n")
cat(sprintf("%.4f mg to %.2f mg\n", predicted_min_tissue, predicted_max_tissue))
cat(rep("=", 60), "\n", sep = "")

# Create formatted values for inline use
max_tissue_formatted <- paste0("`", round(predicted_max_tissue, 2), "mg`")
min_tissue_formatted <- paste0("`", round(predicted_min_tissue, 4), "mg`")
```

Although this is _far_ from an absolute determination, due to the lack of replicates and poor pipetting, we can still use these numbers as a rough guideline to determine whether or not a sample will likely need to be diluted before assaying or can be used directly without dilution.

## SUMMARY

Things we've learned:

1. Pellet insoluble material before pipetting.

2. Glucose does not really appear to be present (detectable) in ctenidia tissue. This suggests that we can skip running this aspect of the assay going forward. However, some additional testing would be beneficial.

3. Roughly, the maximum tissue amount to use undiluted is: `r max_tissue_formatted`.

4. Roughly, the minimum tissue amount to detect glycogen is: `r min_tissue_formatted`.
