---
author: Sam White
toc-title: Contents
toc-depth: 5
toc-location: left
title: sRNA-seq Paired Read Merging Exploratios - Coral E5 Deep Dive
date: '2024-02-02'
draft: true
engine: knitr
categories: 
  - 2024
  - E5
  - sRNA-seq
  - merge
  - bbmerge
  - fastp
---

# INTRO


# SETUP

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = FALSE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  width = 1000         # adds scroll bar
)
```



## Create a Bash variables file

This allows usage of Bash variables across R Markdown chunks.

```{r save-bash-variables-to-rvars-file, engine='bash', eval=TRUE}
{
echo "#### Assign Variables ####"
echo ""

echo "## Data URL"
echo 'export raw_reads_url="https://owl.fish.washington.edu/nightingales/P_evermanni/30-852430235/"'

echo "# Data directories"
echo 'export raw_reads_dir="./raw"'
echo 'export trimmed_fastqs_dir="./trimmed-reads"'
echo ""

echo "## NEB nebnext-small-rna-library-prep-set-for-illumina adapters"
echo 'export first_adapter="AGATCGGAAGAGCACACGTCTGAACTCCAGTCAC"'
echo 'export second_adapter="GATCGTCGGACTGTAGAACTCTGAACGTGTAGATCTCGGTGGTCGCCGTATCATT"'
echo ""

echo "# Input/output files"
echo 'export NEB_adapters_fasta=NEB-adapters.fasta'
echo ""

echo "# Paths to programs"
echo 'export programs_dir="/home/shared"'
echo 'export bbmerge="${programs_dir}/bbmap-39.06/bbmerge.sh"'
echo 'export fastp="${programs_dir}/fastp"'
echo ""

echo "# Set number of CPUs to use"
echo 'export threads=46'
echo ""

echo "# Initialize arrays"
echo 'export trimmed_fastqs_array=()'


} > .bashvars

cat .bashvars
```

## Download raw sRNAseq reads

Reads are downloaded from <https://owl.fish.washington.edu/nightingales/P_evermanni/30-852430235/>

The `--cut-dirs 3` command cuts the preceding directory structure (i.e. `nightingales/P_evermanni/30-852430235/`) so that we just end up with the reads.

```{bash download-raw-reads, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

mkdir --parents "${raw_reads_dir}" "${trimmed_fastqs_dir}"

wget \
--directory-prefix ${raw_reads_dir} \
--recursive \
--no-check-certificate \
--continue \
--cut-dirs 3 \
--no-host-directories \
--no-parent \
--quiet \
--accept "sRNA-POR-79*,checksums.md5" ${raw_reads_url}

ls -lh "${raw_reads_dir}"
```

### Verify raw read checksums

```{bash verify-raw-read-checksums, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

cd "${raw_reads_dir}"

# Checksums file contains other files, so this just looks for the sRNAseq files.
grep "sRNA-POR-79" checksums.md5 | md5sum --check
```

## Create adapters FastA

```{bash create-FastA-of-adapters, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

echo "Creating adapters FastA."
echo ""
adapter_count=0

# Check for adapters file first
# Then create adapters file if doesn't exist
if [ -f "${NEB_adapters_fasta}" ]; then
  echo "${NEB_adapters_fasta} already exists. Nothing to do."
else
  for adapter in "${first_adapter}" "${second_adapter}"
  do
    adapter_count=$((adapter_count + 1))
    printf ">%s\n%s\n" "adapter_${adapter_count}" "${adapter}"
  done >> "${NEB_adapters_fasta}"
fi

echo ""
echo "Adapters FastA:"
echo ""
cat "${NEB_adapters_fasta}"
echo ""
```

## [`fastp`](https://github.com/OpenGene/fastp)

[`fastp`](https://github.com/OpenGene/fastp) has many built-in functionality, including before/after trimming plots, and read 1 and read 2 merging.

### PolyG Trimming Only
```{bash fastp-polyG-only, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

"${fastp}" \
--in1 ./raw/sRNA-POR-79-S1-TP2_R1_001.fastq.gz \
--in2 ./raw/sRNA-POR-79-S1-TP2_R2_001.fastq.gz \
--out1 ./trimmed-reads/fastp-polyG_only.R1.fq.gz \
--out2 ./trimmed-reads/fastp-polyG_only.R2.fq.gz \
--disable_adapter_trimming \
--trim_poly_g \
--overlap_len_require 17 \
--thread ${threads} \
--html "fastp-polyG-only.html" \
--json "fastp-polyG-only.json" \
--report_title "fastp-polyG-only"

```

### Adapters & PolyG Trimming
```{bash fastp-adapters-and-polyg, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

"${fastp}" \
--in1 ./raw/sRNA-POR-79-S1-TP2_R1_001.fastq.gz \
--in2 ./raw/sRNA-POR-79-S1-TP2_R2_001.fastq.gz \
--out1 ./trimmed-reads/fastp-adapters-polyG.R1.fq.gz \
--out2 ./trimmed-reads/fastp-adapters-polyG.R2.fq.gz \
--adapter_fasta NEB-adapters.fasta \
--trim_poly_g \
--overlap_len_require 17 \
--thread ${threads} \
--html "fastp-adapters-polyG.html" \
--json "fastp-adapters-polyG.json" \
--report_title "fastp-adapters-polyG-only"
```

### PolyG Trimming Max Length 31

This can't be run because we know from the polyG only trimming, that most read lengths are >89bp. As such, it would end up discarding all the reads...

### Adapters & PolyG Trimming Max Length 31
```{bash fastp-adapters-and-polyg, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

"${fastp}" \
--in1 ./raw/sRNA-POR-79-S1-TP2_R1_001.fastq.gz \
--in2 ./raw/sRNA-POR-79-S1-TP2_R2_001.fastq.gz \
--out1 ./trimmed-reads/fastp-adapters-polyG.R1.fq.gz \
--out2 ./trimmed-reads/fastp-adapters-polyG.R2.fq.gz \
--adapter_fasta NEB-adapters.fasta \
--trim_poly_g \
--overlap_len_require 17 \
--length_limit 31 \
--thread ${threads} \
--html "fastp-31bp-adapters-polyG.html" \
--json "fastp-31bp-adapters-polyG.json" \
--report_title "fastp-31bp-adapters-polyG-only"
```

### Adapters & PolyG Trimming Max Length 31 With Merge
```{bash fastp-adapters-and-polyg, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

"${fastp}" \
--in1 ./raw/sRNA-POR-79-S1-TP2_R1_001.fastq.gz \
--in2 ./raw/sRNA-POR-79-S1-TP2_R2_001.fastq.gz \
--adapter_fasta NEB-adapters.fasta \
--trim_poly_g \
--overlap_len_require 17 \
--length_limit 31 \
--merge \
--merged_out ./trimmed-reads/fastp-31bp-merged-adapters-polyG.fq.gz \
--thread ${threads} \
--html "fastp-31bp-merge-adapters-polyG.html" \
--json "fastp-31bp-merge-adapters-polyG.json" \
--report_title "fastp-31bp-adapters-polyG-only"
```

### Adapters & PolyG Trimming
```{bash fastp-adapters-and-polyg, engine='bash', eval=TRUE}
# Load bash variables into memory
source .bashvars

"${fastp}" \
--in1 ./raw/sRNA-POR-79-S1-TP2_R1_001.fastq.gz \
--in2 ./raw/sRNA-POR-79-S1-TP2_R2_001.fastq.gz \
--adapter_fasta NEB-adapters.fasta \
--trim_poly_g \
--overlap_len_require 17 \
--merge \
--merged_out ./trimmed-reads/fastp-merged-adapters-polyG.fq.gz \
--thread ${threads} \
--html "fastp-merge-adapters-polyG.html" \
--json "fastp-merge-adapters-polyG.json" \
--report_title "fastp-adapters-polyG-only"
```
