---
author: Sam White
toc-title: Contents
toc-depth: 5
toc-location: left
title: Software Testing - Stevens workflow-annotation Pipeline
date: '2026-03-05'
draft: false
freeze: true
engine: knitr
format:
  html:
    code-fold: show
categories: 
  - "2026"
  - "software"
  - "annotation"
  - "workflow-annotation"
---
# INTRO

Steven asked me to test out his [`workflow-annotation` pipeline](https://github.com/sr320/workflow-annotation/tree/main) (GitHub) after Hannah encountered an issue yesterday and we [implemented a potential fix](https://github.com/RobertsLab/resources/issues/2404#issuecomment-4000199854) (GitHub Issue).

The pipeline was designed to take nucleotide sequences all the way through gene ontology annotation.

After some back-and-forth attempts at troubleshooting, I resorted to asking the Gemini 3 Flash (preview) gAI agent. The suggestion was the lack of an `User-Agent` in the URL request header was causing the 403 permission denied error Hannah (and I) encountered.

I implemented the suggested changes and tested them.

# METHODS

As input, I used a subset of a FastA file provided by Hannah (see code below).

### Set up Conda environment

## Create Mamba/Conda environment

::: {.callout-important}
Be sure to deactivate any current Python/Conda environments!

`deactivate`

and/or

`conda deactivate`
:::

```bash
mamba create \
-n sr320-workflow-annotation \
python=3.11 \
diamond \
blast \
requests \
pandas \
goatools \
matplotlib \
-c bioconda \
-c conda-forge \
-y
```

## R setup

```{r setup, include=TRUE}
library(reticulate)
```

## Set R variables

Specifically, sets Conda paths/environments to be used throughout this notebook.

::: {.callout-important}
This was run on my laptop, so the paths are specific to my system. Adjust as needed for your system.
:::

```{r R-variables, eval=TRUE}
# CONDA
conda_env_name <- c("/home/sam/programs/miniforge3/envs/sr320-workflow-annotation")
conda_path <- c("/home/sam/programs/miniforge3/bin/conda")

# WORKING DIRECTORY
working_dir <- c("/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/")
Sys.setenv(working_dir = working_dir)
```

## Load [worfkflow-annotation](https://github.com/sr320/workflow-annotation/tree/main) conda environment

If this is successful, the first line of output should show that the Python being used is the one in your [workflow-annotation](https://github.com/sr320/workflow-annotation/tree/main) conda environment path.

E.g.

`python:         /home/sam/programs/miniforge3/envs/sr320-workflow-annotation/bin/python`

```{r load-conda-env, eval=TRUE}
use_condaenv(condaenv = conda_env_name, conda = conda_path)
py_config()
```


### Clone the workflow-annotation pipeline

```bash
git clone git@github.com:kubu4/workflow-annotation.git
```

### Download the FastA

::: {.callout-note}
The full FastA was not utilized. I created a subset for quick testing.
:::

```bash
cd "/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/"

git clone git@github.com:kubu4/workflow-annotation.git

cd workflow-annotation

wget --quiet --continue https://raw.githubusercontent.com/hannahnowers/Seastar-capstone/refs/heads/main/data/derm-protein.fa
```

```{bash peek-fasta-subset}
cd "/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/workflow-annotation"

cat subset-derm-protein.fa
```

### Execute the annotation script


```{bash annotate}
cd "/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/workflow-annotation" && \
export PATH="/home/sam/programs/miniforge3/envs/sr320-workflow-annotation/bin:$PATH" && \
./blast2slim.sh \
-i subset-derm-protein.fa \
--diamond \
--protein \
> blast_run.log 2>&1
```

# RESULTS

### Check log
```{bash check-log}
cat "/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/workflow-annotation/blast_run.log"
```

### Confirm expected output files
```{bash confirm-output-files}
cd "/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/workflow-annotation"

ls -ltrh output/run_*
```


### Peek at output files
```{bash peek-output-files}
cd "/home/sam/gitrepos/RobertsLab/sams-notebook/posts/2026/2026-03-05-Software-Testing---Stevens-workflow-annotation-Pipeline/workflow-annotation/output"

find run_*/ -maxdepth 1 -type f ! -name "*.png" -exec head {} +
```

# DISCUSSION

Looks like the fix worked! Everything is as expected.



