---
author: Sam White
toc-title: Contents
toc-depth: 5
toc-location: left
title: Software Testing - Stevens workflow-annotation Pipeline
date: '2026-03-05'
draft: true
engine: knitr
categories: 
  - "2026"
  - "software"
  - "annotation"
  - "workflow-annotation"
---
# INTRO

Steven asked me to test out his [`workflow-annoation` pipeline](https://github.com/sr320/workflow-annotation/tree/main) (GitHub) after Hannah encountered an issue yesterday and we [implemented a potential fix](https://github.com/RobertsLab/resources/issues/2404#issuecomment-4000199854) (GitHub Issue).

The pipeline was designed to take nucleotide sequences all the way through gene ontology annotation. As input, I used a small FastA file provided by Hannah (see code below).

## METHODS

### Set up Conda environment

## Create Mamba/Conda environment

::: {.callout-important}
Be sure to deactivate any current Python/Conda environments!

`deactivate`

and/or

`conda deactivate`
:::

```bash
mamba create \
-n sr320-workflow-annotation \
python=3.11 \
diamond \
requests \
pandas \
goatools \
matplotlib \
-c bioconda \
-c conda-forge \
-y
```

## R setup

```{r setup, include=TRUE}
library(reticulate)
```

## Set R variables

Specifically, sets Conda paths/environments to be used throughout this notebook.
```{r R-variables, eval=TRUE}
# CONDA
conda_env_name <- c("/home/sam/programs/mambaforge/envs/sr320-workflow-annotation")
conda_path <- c("/opt/anaconda/anaconda3/bin/conda")
```

## Load [worfkflow-annotation](https://github.com/sr320/workflow-annotation/tree/main) conda environment

If this is successful, the first line of output should show that the Python being used is the one in your [worfkflow-annotation](https://github.com/sr320/workflow-annotation/tree/main) conda environment path.

E.g.

`python:         /home/sam/programs/mambaforge/envs/sr320-workflow-annotation/bin/python`

```{r load-barnacle-conda-env, eval=TRUE}
use_condaenv(condaenv = conda_env_name, conda = conda_path)
py_config()
```


### Clone the workflow-annotation pipeline

```{r download-pipeline, engine='bash'}
git clone git@github.com:sr320/workflow-annotation.git
```

### Download the FastA
```{r download-fasta, engine='bash'}
cd workflow-annotation

wget --quiet --continue https://raw.githubusercontent.com/hannahnowers/Seastar-capstone/refs/heads/main/data/derm-protein.fa

ls -ltrh
```

### Execute the annotation script

```{r annotate, engine='bash'}
cd workflow-annotation

bash blast2slim.sh \
-i derm-protein.fa
--protein
> blast_run.log 2>&1

```
